---
title: "Hands-on-Exercise 1"
author: "Marcus Jaeson Yeo"
---

# Getting Started

### 1.0 Install and launch R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, tidyverse packages will be launched into R.

```{r}
pacman::p_load(sf, tidyverse)
```

## 2.0 Importing Data (Geospatial)

### (MasterPlan) Importing shapefile format files as polygon feature dataframe

```{r}
mpsz = st_read(dsn = "data/geospatial/MasterPlan2014SubzoneBoundaryWebSHP", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

### (Cycling Path) Importing shapefile format files as polygon feature dataframe

```{r}
cycling_path = st_read(dsn = "data/geospatial/CyclingPath_Jul2024/CyclingPath_Jul2024", 
                  layer = "CyclingPathGazette")
```

### (Preschool) Importing GIS data in kml format

```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

## 3.0 Checking content of data frames (Simple Feature)

-   st_geometry()

```{r}
st_geometry(mpsz)
```

-   glimpse()

```{r}
glimpse(mpsz)
```

-   head()

```{r}
head(mpsz, n=5)
```

## 4.0 Plotting data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## 5.0 Projection Transformation

```{r}
st_crs(mpsz)
```

Assigning the correct EPSG code to mpsz

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

Double-check the EPSG code of mpsz

```{r}
st_crs(mpsz3414)
```

5.1 Transforming the projection of preschool from wgs84 to svy21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
st_geometry(preschool3414)
```

## 6.0 Importing data (Aspatial)

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

### Listing the column names

```{r}
colnames(listings)
```

### Plotting a simple bar chart

let's pick **neighbourhood** group

```{r}

ggplot(data = listings, aes(x = neighbourhood_group)) + geom_bar()
```

Converting aspatial dataframe into simple feature dataframe

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)

glimpse(listings_sf)
```

## 7.0 Cycling Path Scenario

1.  Computing 5 meter buffers around cycling paths

```{r}
buffer_cycling <- st_buffer(cycling_path, 
                               dist=5, nQuadSegs = 30)
```

2.  calculate the area around the buffers

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

3.  Find total land involved

```{r}
sum(buffer_cycling$AREA)
```

## 8.0 Preschool Scenario

1.  Identify number of preschools in **each** planning subzone

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

2.  Summary statistics of **preschool count**

```{r}
summary(mpsz3414$`PreSch Count`)
```

3.  List the planning subzone with **most** number of preschools

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

4.  Calculating density of preschool by planning subzone

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()

mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## 9.0 EDA

1.  Histogram of distribution of PreSch Density

```{r}
hist(mpsz3414$`PreSch Density`)
```

2.  Geometric Histogram of Distribution of preschools

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

3.  Scatterplot showing relationship between preschool count vs preschool Density

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
