---
title: "In-class Exercise 04: Forest Fire in Kepulauan Bangka Belitung, Indonesia (Spatio-temporal Point Patterns Analysis"
author: "Marcus Jaeson Yeo"
date: "September 09, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
---

## 1.0 Introduction

### 1.1 Forest Fire in Kepulauan Bangka Belitung, Indonesia

### 1.2 What are we trying to answer?

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

------------------------------------------------------------------------

## 2.0 Importing Packages

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

------------------------------------------------------------------------

## 3.0 Importing Datasets into Environment

-   importing kbb shapefiles (geospatial)

```{r}
kbb_sf <- st_read(dsn = "data/rawdata", 
                     layer="Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

-   Importing and Preparing Forest Fire data (aspatial)

    ```{r}
    fire <- read_csv("data/rawdata/forestfires.csv")
    ```

    ```{r}
    fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
      st_as_sf(coords = c("longitude", "latitude"),
               crs = 4326) %>%
      st_transform(crs=32748)
    ```

    ------------------------------------------------------------------------

------------------------------------------------------------------------

## 4.0 owin Object Transformation

-   convert to owin (as.owin())

    ```{r}
    kbb_owin <- as.owin(kbb_sf)
    kbb_owin
    ```

    ```{r}
    class(kbb_owin)
    ```

------------------------------------------------------------------------

## 5.0 Data Wrangling

### 5.1 mutate

-   why do we need to convert (mutate) data type of *acq_date* to numeric?

    -   because ppp object only accept numerical or character as mark

```{r}
fire_sf_mutated_date <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, 
                           label = TRUE,
                           abbr= FALSE))

```

### 5.2 Plot / VIsualising spatial points

-   Visualising geographic distribution of forest fire points all throughout

    ```{r}
    tm_shape(kbb_sf) +   
      tm_polygons() +    
        tm_shape(fire_sf_mutated_date) +    
          tm_dots()
    ```

    `{r}`

-   Visualising geographic distribution of forest fire points by month

    ```{r}
    tm_shape(kbb_sf) +   
      tm_polygons() +    
        tm_shape(fire_sf_mutated_date) +    
          tm_dots(size = 0.1) +    
            tm_facets(by="Month_fac",              
                      free.coords = FALSE,             
                      drop.units = TRUE)
    ```

### 5.3 computing STKDE by Month (time series KDE)

```{r}
fire_month <- fire_sf_mutated_date %>%   
  select(Month_num)
```

```{r}
fire_month_ppp <- as.ppp(fire_month) fire_month_ppp
```

### 5.4 summary + check for duplicates

```{r}
summary(fire_month_ppp)
```

```{r}
any(duplicated(fire_month_ppp))
```

### 5.5 include created owin object

-   combining origin_am_ppp and am_owin objects

    -   fire_month_ppp + kbb_owin

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

```{r}
plot(fire_month_owin)
```

------------------------------------------------------------------------

## 6.0 Compute Spatio-temporal KDE

-   use spattemp.density() from sparr package, to compute the STKDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

------------------------------------------------------------------------

## 7.0 Plotting Spatio-temporal KDE object

-   plots KDE between July 2023 - December 2023

```{r}
tims <- c(7,8,9,10,11,12) 
par(mfcol=c(2,3)) 
for(i in tims) {
  plot(st_kde, 1,         
  override.par=FALSE,        
  fix.range=TRUE,        
  main=paste("KDE at month", 1)) 
}
```

------------------------------------------------------------------------

## 8.0 Computing STKDE by [Day of Year]{.underline}

### 8.1 Create ppp object 

```{r}
fire_yday_ppp <- fire_sf_mutated_date %>% 
  select(DayofYear) %>%
  as.ppp()

```

### 8.2 Combine both ppp object + owin object (kbb) = combined owin object 

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

### 8.3 Compute STKDE by day of Year

```{r}
kde_yday <- spattemp.density(fire_yday_owin)
summary(kde_yday)
```

------------------------------------------------------------------------

## 9.0 Plot STKDE by day of Year

-   temporal bound: t=10 to 352

```{r}
plot(kde_yday)
```

## References

-   tmap: <https://r-tmap.github.io/tmap/reference/index.html>

-   sparr: <https://cran.r-project.org/web/packages/sparr/index.html>
