---
title: "Take-home Exercise 2: Application of Geospatial Analysis Methods to Discover Thailand Drug Abuse at the Province Level"
author: "Marcus Jaeson Yeo"
date: "September 26, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
---

## 1.0 Introduction

#### Context:

Drug abuse is associated with significant negative health, financial and social consequences. Yet, illicit drug consumption remains highly prevalent and continues to be a growing problem worldwide. In 2021, 1 in 17 people, aged 15–64 in the world had used a drug in the past 12 months. Notwithstanding population growth, the estimated number of drug users grew from 240 million in 2011 to 296 million in 2021.

The geopolitics of Thailand which is near the [Golden Triangle](https://en.wikipedia.org/wiki/Golden_Triangle_(Southeast_Asia)) of Indochina, the largest drug production site in Asia, and the constant transportation infrastructure development made Thailand became market and transit routes for drug trafficking to the third countries.

In Thailand, drug abuse is one of the major social issue. There are about 2.7 million youths using drugs in Thailand. Among youths aged between 15 and 19 years, there are about 300,000 who have needs for drug treatment. Most of Thai youths involved with drugs are vocational-school students, which nearly doubles in number compared to secondary-school students.

Figure below shows geographic distribution of drug use cases by province and by year.

#### Objectives:

-   if the key indicators of drug abuse of Thailand are independent from space. (spatial independent, randomly distributed) (spatial weights and applications)

-   If the indicators of drug abuse is indeed spatial dependent, then, you would like to detect where are the clusters and outliers, and the hotspots. (no need for emerging hotspots)

-   Last but not least, you are also interested to investigate how the observation above evolve over time. (no need for spatial-temporal)

------------------------------------------------------------------------

## 2.0 Importing Packages

```{r}
pacman::p_load(sf, lubridate, raster, spatstat, sparr, tmap, tidyverse, spdep, sfdep, leaflet, smoothr)
```

------------------------------------------------------------------------

## 3.0 Importing Datasets into Environment

-   geospatial (admin1 (province level))

    ```{r}
    thailand_boundary = st_read(dsn="data/geospatial/",
            layer="tha_admbnda_adm1_rtsd_20220121")

    dim(thailand_boundary)
    ```

    ```{r}
    # plot(thailand_boundary, max.plot=16)
    ```

-   aspatial

    ```{r}
    thailand_sf <- read_csv("data/aspatial/thai_drug_offenses_2017_2022.csv")

    dim(thailand_sf)
    ```

    ```{r}
    # plot(thailand_sf)
    ```

------------------------------------------------------------------------

## 4.0 Data Wrangling + necessary Projection Transformation

### 4.1 Dropping Redundant Columns

-   Dropping columns with NA and columns that do not give us valuable information

```{r}
thailand_boundary <- thailand_boundary %>%
  select(1:3, 14:17)

thailand_sf <- thailand_sf %>%
  select(1:3, 5)
```

#### 4.2 Projection Transformation

-   We need to further understand a bit more about the common identifier (province), hence printing the unique provinces, which is both [76]{.underline} unique provinces for both boundary and sf.

```{r}
unique_provinces_boundary <- unique(thailand_boundary$ADM1_EN)
unique_provinces_boundary
```

```{r}
unique_provinces_sf <- unique(thailand_sf$province_en)
unique_provinces_sf
```

-   After studying the columns of both the geospatial and aspatial data, we see that the aspatial data (thailand_sf) lacks geometrical data (such as the latitude and longitude spatial features). We need to join both boundary and working data with its common columns: **ADM1_EN** (thailand_boundary) and **province_en** (thailand_sf).

    -   First, we need to mutate to get the common identifier

```{r}
#| eval: false
thailand_boundary <- thailand_boundary %>%
  mutate(ADM1_EN = str_to_lower(ADM1_EN),  # Convert to lowercase
         ADM1_EN = str_trim(ADM1_EN))

thailand_sf <- thailand_sf %>%
  mutate(province_en = str_to_lower(province_en),  # Convert to lowercase
         province_en = str_trim(province_en))
         
```

-   Second, we perform a left join

```{r}
#| eval: false
thailand <- thailand_boundary %>%
  left_join(thailand_sf, by = c("ADM1_EN" = "province_en")) %>%
  select(3:4, 7:10)
```

Since the EPSG code is 4326, which is the global code. We want to change both the EPSG code of the boundary data and working data to Thailand's global projection system. Acquiring the code from <https://epsg.io/32647>

-   Third, we transform the it to Thailand's projected coordinate system (UTM Zone 47N) with the EPSG code: 32647

```{r}
#| eval: false
thailand <- thailand %>%
  st_transform(crs=32647)
```

```{r}
#| eval: false
st_crs(thailand)
```

#### 4.3 Plotting the object to visualise structure

```{r}
#| eval: false
thailand_shp = st_read("data/geospatial/tha_admbnda_adm1_rtsd_20220121.shp")
```

```{r}
#| eval: false
boundary_plot <- tmap_mode("plot")
tm_shape(thailand_shp)+
  tm_fill(col="white")+
  tm_borders(col = "black", lwd=0.3, alpha=0.6)+
  tm_layout(
    main.title = "Provinces",
    main.title.size = 1,
    main.title.position = "center",
    legend.show = FALSE,
     frame = FALSE)
```

```{r}
#| eval: false
# basemap <- tm_shape(thailand) +
#   tm_polygons() +
#   tm_text("ADM1_EN", size=0.5)
# 
# drug_cases <- qtm(thailand, "no_cases")
# tmap_arrange(basemap, drug_cases, asp=1, ncol=2)
```

#### 4.4 Writing and Reading into RDS format

```{r}
#| eval: false

#thailand combined dataframe
write_rds(thailand, "data/rds/thailand.rds")

#plot
write_rds(boundary_plot, "data/rds/boundary_plot.rds")
# write_rds(basemap, "data/rds/basemap.rds")
# write_rds(drug_cases, "data/rds/drug_cases.rds")
```

```{r}

thailand <- read_rds("data/rds/thailand.rds")
```

------------------------------------------------------------------------

## 5.0 Exploratory Spatial Data Analysis

### 5.1 Further data wrangling and shortening

### 5.1.1 Dropping and summing total number of drug cases

-   Lets drop both date and types_of_drug_offenses column

```{r}
colnames(thailand)
```

```{r}
thailand <- thailand %>%
  select("ADM1_EN", "fiscal_year", "no_cases", "geometry")

thailand
```

-   Sum all the no_cases by each year and province and sort them ascending, drop rows where the total_cases are 0

```{r}
thailand <- thailand %>%
  group_by(fiscal_year, ADM1_EN) %>%
  summarize(total_cases = sum(no_cases, na.rm = TRUE)) %>%
  filter(total_cases > 0)  # Drop rows where total_cases = 0
```

### 5.1.2 Understanding our data

-   unique years

```{r}
unique_years = unique(thailand$fiscal_year)
unique_provinces = unique(thailand$ADM1_EN)

unique_years
cat('\n')
unique_provinces
```

-   min and max total no of drug cases

```{r}
min_cases = min(thailand$no_cases, na.rm = FALSE)
max_cases = max(thailand$no_cases, na.rm = FALSE)

min_cases
cat("\n")
max_cases
```

### 5.2 Plotting

```{r}
# Ensure geometry is retained for mapping
thailand_summary <- st_as_sf(thailand)
```

### 5.2.1 Total number of cases by each year and each province

```{r}
# Create the faceted plot for all years with desired layout
faceted_plot <- tm_shape(thailand_summary) +
  tm_fill("total_cases", 
          n = 5, 
          style = "equal", 
          palette = "Blues", 
          title = "Drug use cases") +
  tm_borders(alpha = 0.5) +
  tm_facets(by = "fiscal_year", 
            ncol = 2,               # Ensure side-by-side layout with 2 columns
            drop.empty.facets = TRUE) +  # Drop missing facets (empty years)
  tm_layout(main.title = "Drug Use Cases by Year", 
            title.size = 1.2,
            legend.outside = TRUE,   # Legend outside the map area
            asp = 1,                 # Control the aspect ratio for larger map
            frame = FALSE,           # Removes white border frame around maps
            inner.margins = c(0.02, 0.02, 0.02, 0.02))  # Reduce white space

# Print the faceted plot
faceted_plot
```

### 5.2.2 Side by side view of province and total no of drug cases in year 2017

```{r}

provinces <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) +
  tm_polygons() +
  tm_text("ADM1_EN", size=0.5)

custom_breaks <- c(0, 5000, 10000, 15000, 20000)  # Customize as necessary

# total_cases <- qtm(thailand_summary, "total_cases")
total_cases <- tm_shape(thailand_summary) +
  tm_polygons("total_cases", 
              breaks = custom_breaks,  # Apply the custom breaks
              palette = "Blues",        # Color palette
              title = "Total Cases")    # Title for the legend
tmap_arrange(provinces, total_cases, asp=1, ncol=2)

```

### 5.2.3 Top 5 Provinces with the highest number of drug cases in year 2017

```{r}
# Step 1: Find the top 5 provinces with the highest total cases
top_5_provinces <- thailand_summary %>%
  filter(fiscal_year == 2017) %>%
  arrange(desc(total_cases)) %>%
  slice_head(n = 5)  # Select the top 5 rows

# Step 2: Highlight the top 5 provinces
provinces <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) +
  tm_polygons() +
  tm_text("ADM1_EN", size = 0.5)

# Apply custom breaks to the total_cases plot
total_cases <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) +
  tm_polygons("total_cases", 
              breaks = custom_breaks,  # Apply the custom breaks
              palette = "Blues",        # Color palette
              title = "Total Cases") +
  tm_borders(alpha = 0.5)

# Step 3: Highlight the top 5 provinces on the same plot
highlight_top_5 <- tm_shape(top_5_provinces) +
  tm_borders(col = "red", lwd = 3) +  # Use red borders to highlight
  tm_text("ADM1_EN", size = 0.7, col = "red", fontface = "bold")  # Optional: Label them

# Step 4: Arrange the plots
tmap_arrange(provinces, total_cases + highlight_top_5, asp = 1, ncol = 2)
```

```{r}
#|eval: false
# plot_2018 <- tm_shape(thailand_summary %>% filter(fiscal_year == 2018)) + 
#   tm_fill("total_cases", n = 5, style = "equal") + tm_borders() + tm_layout(main.title = "2018")
# 
# plot_2019 <- tm_shape(thailand_summary %>% filter(fiscal_year == 2019)) + 
#   tm_fill("total_cases", n = 5, style = "equal") + tm_borders() + tm_layout(main.title = "2019")
# 
# plot_2020 <- tm_shape(thailand_summary %>% filter(fiscal_year == 2020)) + 
#   tm_fill("total_cases", n = 5, style = "equal") + tm_borders() + tm_layout(main.title = "2020")
# 
# plot_2021 <- tm_shape(thailand_summary %>% filter(fiscal_year == 2021)) + 
#   tm_fill("total_cases", n = 5, style = "equal") + tm_borders() + tm_layout(main.title = "2021")
# 
# plot_2022 <- tm_shape(thailand_summary %>% filter(fiscal_year == 2022)) + 
#   tm_fill("total_cases", n = 5, style = "equal") + tm_borders() + tm_layout(main.title = "2022")
# 
# # Arrange them
# tmap_arrange(plot_2017, plot_2018, plot_2019, plot_2020, plot_2021, plot_2022, ncol = 2)
```

### 5.3 Visualising Equal and Quantile in year 2017

```{r}
equal <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) + 
  tm_polygons("total_cases",
          palette = "Blues",
          style="equal", n=10) + 
  tm_borders(alpha = 0.5) + 
  tm_layout(main.title = "Equal interval classification (year 2017)", title.size = 2)

# quantile <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) +
#   tm_fill("total_cases",
#           n = 5,
#           style = "quantile") +
#   tm_borders(alpha = 0.5) +
#   tm_layout(main.title = "Equal quantile classification")

quantile <- tm_shape(thailand_summary %>% filter(fiscal_year == 2017)) +
  tm_polygons("total_cases",
          palette = "Blues",
          style="quantile", n=10) +
  tm_layout(main.title = "Quantile interval classification (year 2017)", title.size = 2)

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

```{r}
write_rds(thailand, "data/rds/thailand.rds")
write_rds(equal, "data/rds/equal.rds")

write_rds(quantile, "data/rds/quantile.rds")
```

------------------------------------------------------------------------

## 6.0 Spatial Weights and Applications

------------------------------------------------------------------------

## 7.0 Global Measures of Spatial Autocorrelation

```{r}
#| eval: false
wm_q <- poly2nb(thailand, 
                queen = TRUE)

summary(wm_q)
```

```{r}
#| eval: false
write_rds(wm_q, "data/rds/wm_q.rds")
```

```{r}
wm_q <- read_rds("data/rds/wm_q.rds")
```

### 7.1 Calculating the neighbours and weights

```{r}
# wm_q.nb <- st_contiguity(thailand, queen=TRUE)
# wm_q.wt <- st_weights(wm_q.nb, style = "W")
# wm_q.count <- thailand$count
```

### 7.2 Construct spatial contiguity matrices (weights) and row-standardised weight matrices

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

### 7.3 Global Measures of Spatial Autocorrelation: Moran's I test

### 7.3.1 Moran's I Test

```{r}
moran.test(thailand$total_cases, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

### 7.3.2 Computing Monte Carlo Moran's I

-   set a seed of 1234

-   number of simulations = 1000

```{r}
set.seed(1234)
bperm_thailand= moran.mc(thailand$total_cases, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm_thailand
```

### 7.3.3 Visualising Monte Carlo Moran's I

-   Computing some basic statistics

```{r}
#compute mean
mean(bperm_thailand$res[1:999])
cat('\n')

#compute variance
var(bperm_thailand$res[1:999])
cat('\n')

#summary bperm
summary(bperm_thailand$res[1:999])
```

-   plot basic histogram using base graph

```{r}
hist(bperm_thailand$res, #x
     freq=TRUE, #y
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

-   plotting histogram using ggplot2

```{r}
data <- data.frame(simulated_morans_i = bperm_thailand$res[1:999])

ggplot(data, aes(x = simulated_morans_i)) +
  geom_histogram(binwidth = (max(data$simulated_morans_i) - min(data$simulated_morans_i)) / 20, 
                 color = "black", fill = "pink") +  # Histogram
  geom_vline(xintercept = 0, color = "red", linetype = "solid") +  # Vertical line at x=0
  labs(x = "Simulated Moran's I", y = "Frequency") +  # Labels for x and y axis
  theme_minimal()  # Clean minimal theme

```

------------------------------------------------------------------------

## 8.0 Local Measures of Spatial Autocorrelation

------------------------------------------------------------------------

## 10.0 Conclusion

------------------------------------------------------------------------

## References
